---
title: "Realtor.com - Housing Market in California"
author: "Alondra Aleman, Blessing Nwogu, Stephanie Garcia"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
---



```{r setup, include=FALSE}
# install.packages("flexdashboard")
# install.packages("dplyr")
# install.packages("plotly")
# install.packages("DT")
# install.packages("gridExtra")
# install.packages("shiny")
# install.packages("tidyverse")
# install.packages("ggplot2")
# install.packages("htmltools")
# install.packages("highcharter")
# install.packages("viridis")

library(flexdashboard)
library(dplyr)
library(plotly)
library(DT)
library(gridExtra)
library(shiny)
library(tidyverse)
library(ggplot2)
library(htmltools)
library(highcharter)
library(viridis)
```




```{r include=FALSE}
#importing the dataset, selecting variables needed and removing null values
#read csv file, select variables and clean data
realtor_data = read.csv("Realtor_data.csv") #Loading California Realtor.com data
dates = as.Date(realtor_data$DATE, format = "%m/%d/%y")
sum(is.na(realtor_data)) #check for missing data
summary(realtor_data)

hotness = read.csv("RDC_CA_Hotness.csv")
str(hotness)
summary(hotness)

df = read.csv("RDC_Inventory_Hotness_Metrics_County_History.csv") 
df = df %>%
  select(year, month, county_name, state, hotness_rank, hotness_score, supply_score, demand_score)

hotness2 = na.omit(df) #remove NAs
str(hotness2)
summary(hotness2)
```


```{r echo=FALSE}
Lowest_prices_2020 =
  hotness %>%
  select(ï..DATE,median_listing_price,COUNTY_NAME,median_days_on_market,hotness_rank,hotness_score,demand_score, supply_score) %>%
  filter(COUNTY_NAME %in% c("fresno","imperial","kings","kern", "lake","madera","merced","tehama","tulare","siskiyou")) %>%
  filter(ï..DATE %in% c("1/1/20","2/1/20","3/1/20","4/1/20","5/1/20","6/1/20","7/1/20","8/1/20","9/1/20","10/1/20","11/1/20","12/1/20"))

#The Hotness score is an equally-weighted composite metric of a geography’s supply score and demand score.
```




```{r echo=FALSE}
prices_2020 =
  hotness %>%
  select(ï..DATE,median_listing_price,COUNTY_NAME, median_days_on_market,hotness_rank,hotness_score,demand_score, supply_score) %>%
  filter(COUNTY_NAME %in% c("santa barbara","san mateo","marin","san francisco","napa","monterey","santa clara","santa cruz","los angeles","orange")) %>%
  filter(ï..DATE %in% c("1/1/20","2/1/20","3/1/20","4/1/20","5/1/20","6/1/20","7/1/20","8/1/20","9/1/20","10/1/20","11/1/20","12/1/20"))
```

Listing Prices
=======================================================================


Row {data-width=350}
-----------------------------------------------------------------------


### Maximum Listing Price in 2020

```{r}
#Value box for the minimum listing prices in 2020
valueBox(scales::dollar(1596134.45), color = "purple", icon = "fa-chart-line")
```

### Average Listing Price in 2020

```{r}
valueBox(scales::dollar(1397441), color = "blue", icon = "fa-hand-holding-usd")
```

### Minimum Listing Price in 2020

```{r}
#Value box for minimum listing price
valueBox(scales::dollar(1218316.55), color = "green", icon = "fa-money-bill")
```

Column {.tabset data-width=650}
-----------------------------------------------------------------------


### Housing Demand
 
```{r}
demandScore_state = hotness2%>%
  group_by(year, state)%>%
  summarise(state, Average = round(mean(demand_score),2)) #group the data by tear and then by state and summarize it by average demand score of states
#view(p)
states= read.csv("US states.csv") #read csv file for US states
a = demandScore_state%>%
  inner_join(states, by.x = state, by.y = state)%>% #join the column 'state' in 'states' dataset to the demandScore_state view 
  select(year,state,state_code, Average)%>%
  mutate(hover = paste0(state, "\n", "Demand Score: ",Average)) #concatenate the hover text
fontStyle = list(family = "DM Sans",
                 size = 15,
                 color ="black"
                 ) #font style
label = list(bgcolor = "#EEEEEE",
             bordercolr = "transparent",
             font = fontStyle
             )
```



```{r}
plotMap= plot_geo(a, 
            locationmode = 'USA-states', 
            frame = ~year)%>% #use USA-States map and add a slider with years in the dataset
  add_trace(locations = ~state_code, 
            z = ~Average,
            zmin = 0,
            zmax = max(a$Average),
            colorscale = 'Picnic', reversescale = T,
            text = ~hover,
            hoverinfo = 'text',
            color = ~Average)%>%
  layout(geo = list(scope = "usa"),
         font = list(family= "DM Sans"),
         title = "Housing Demand\n2017 - 2021")%>%
  style(hoverlabel = label)%>%
  config(displayModeBar = F)
plotMap

```



```{r include=FALSE}
Dec_hotness =
  hotness %>%
  select(ï..DATE,median_listing_price,COUNTY_NAME, median_days_on_market) %>%
  filter(COUNTY_NAME %in% c("santa barbara","san mateo","marin","san francisco","napa","monterey","santa clara","santa cruz","los angeles","orange")) %>%
  filter(ï..DATE %in% "12/1/20")

```

```{r}
Lowest_price =
  hotness %>%
  select(ï..DATE,median_listing_price,COUNTY_NAME) %>%
  filter(COUNTY_NAME %in% c("fresno","imperial","kings","kern", "lake","madera","merced","tehama","tulare","siskiyou")) %>%
  filter(ï..DATE %in% "12/1/20")

```

### 10 Most Expensive Counties in California

```{r fig.height=4, fig.width=7}
fig2 = plot_ly(
  x = Dec_hotness$COUNTY_NAME,
  y = Dec_hotness$median_listing_price,
  name = "Top Highest Prices per county",
  type = "bar",
  marker = list(color = "rgba('195, 195, 195, 0.6)",
                            line = list(color = "rgb(20, 20, 20)",
                                        width = 2 )))
 
fig2 <- fig2 %>% layout(title = "Median Listing Prices in \nTop 10 Most Expensive Counties (Dec 2020)", yaxis = list(title="Listing Price", color ="black"))


fig2
```

### 10 Least Expensive Counties in California

```{r}
fig3 = plot_ly(
  x = Lowest_price$COUNTY_NAME,
  y = Lowest_price$median_listing_price,
  name = "Lowest Prices per county",
  type = "bar",
  marker = list(color = "rgba(255, 0, 0, 0.6)",
                            line = list(color = "rgb(20, 20, 20)",
                                        width = 2))
)
 fig3 <- fig3 %>% layout(title = "Median Listing Prices in \nLeast 10 Expensive Counties (Dec 2020)", yaxis = list(title=" Listing Price", color ="black"))

 fig3
```


Days on the Market
=======================================================================

Column {.tabset data-width=650}
-----------------------------------------------------------------------

### 10 Most Expensive Properties

```{r}
plot_ly(prices_2020, x = ~median_days_on_market, y = ~median_listing_price) %>%
  layout(title = "Listing Price and Days on Market Top 10 Expensive")
```

### 10 Least Expensive Properties
```{r}
plot_ly(Lowest_prices_2020, x = ~median_days_on_market, y = ~median_listing_price) %>%
  layout(title = "Listing Price and Days on Market 10 Least Expensive")
```


Column {.tabset data-width=350}
-----------------------------------------------------------------------
### Distribution of Median Days on Market in the 10 Most Expensive Counties

```{r}
plot_ly() %>%
  add_trace(data = prices_2020,
            y = ~median_days_on_market,
            color = ~COUNTY_NAME,
            type = "box") %>%
  layout(title = "Median Days on Market for the 10 Most Expensive Counties",
         yaxis = list(title = "Days on Market"), xaxis = list(title="Counties"))
```

### Distribution of Median Days on Market in the 10 Least Expensive Counties

```{r}
plot_ly() %>%
  add_trace(data = Lowest_prices_2020,
            y = ~median_days_on_market,
            color = ~COUNTY_NAME,
            type = "box") %>%
  layout(title = "Sales per city",
         xaxis = list(title = "City")) %>%
  layout(title = "Median Days on Market for the 10 Least Expensive Counties",
         yaxis = list(title = "Days on Market"), xaxis = list(title="Counties"))
```


Demand Hotness
=======================================================================

Column {.tabset data-width=650}
-----------------------------------------------------------------------

### 10 Most Expensive Counties


```{r}
plot_ly(prices_2020, x = ~median_listing_price, y = ~hotness_score) %>%
  layout(title = "Hotness Score in the 10 Most Expensive Counties")
```



### 10 Least Expensive Counties



```{r}
plot_ly(Lowest_prices_2020, x = ~median_listing_price, y = ~hotness_score) %>%
  layout(title = "Hotness Score in the 10 Least Expensive Counties")

```


Column {.tabset data-width=350}
-----------------------------------------------------------------------

### 10 Most Expensive Counties

```{r}
plot_ly() %>%
  add_trace(data = prices_2020,
            y = ~hotness_score,
            color = ~COUNTY_NAME,
            type = "box") %>%
  layout(title = "Hotness Score for the 10 Most Expensive Counties ")
```

### 10 Least Expensive Counties

```{r}
plot_ly() %>%
  add_trace(data = Lowest_prices_2020,
            y = ~hotness_score,
            color = ~COUNTY_NAME,
            type = "box") %>%
  layout(title = "Hotness Score in the 10 Least Expensive Counties")
```



